---
title: "STAT545-hw04-An-Byeongchan"
output:
  html_document:
    keep_md: yes

---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(fig.width=4, fig.height=3)
```

```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(gapminder))
```


### Choose your own adventure

- Pick a join prompt and do it.
I used the dataset `areas` which was dealt with in class. I tried to combine gapminder data in 2007 with `areas`. As I used right_join(), `g1` preserves all the rows of `areas`. Here, I did not specify a character vector of variables to join by using `by = "country"`. This was because there was only one variable in common. 
In addition, I compared `full_join()` with `left_join()`. Because `gapminder` does not have `"Vatican City"` which is included in `areas`, `full_join()` creates one more observation (which is for `Vatican City`), compared to `left_join()`
```{r}
areas <- data.frame(country=c("Canada", "United States", "India", "Vatican City"),
                     area=c(998.5*10^6, 983.4*10^6, 328.7*10^6, 44))

g1 <- gapminder %>% 
  filter(year == 2007) %>% 
  right_join(areas)
g1
# The number of observations using full_join()
nrow(full_join(gapminder, areas, by = "country"))
# The number of observations using left_join()
nrow(left_join(gapminder, areas, by = "country"))
```

- Pick one of the data reshaping prompts and do it.
I trid to use `gather()` for long format and `spread()` for wide format.
```{r}
g2 <- g1 %>% 
  select(country, continent, pop, gdpPercap, area) %>% 
  gather(key = "Measures", value = "My_value", pop, gdpPercap, area)
g2
g2 %>% 
  spread(key = "Measures", value = "My_value")
```


### General data reshaping and relationship to aggregation

__Problem__: You have data in one "shape" but you wish it were in another. Usually this is because the alternative shape is superior for presenting a table, making a figure, or doing aggregation and statistical analysis.

__Solution__: Reshape your data. For simple reshaping, `gather()` and `spread()` from `tidyr` will suffice. Do the thing that it possible / easier now that your data has a new shape.

__Prompts__:

Activity #1
- two `tidyr` functions
1. `gather(key = "key", value = "value", ...)`

2. `spread(key = "key", value = "value", ...)`
Make sure that there is no duplicate to use `spread()`. When there are more-than-one observations (which have same values except for "key" values) for the same "key" value, the function does not work. This can be fixed by distinguishing each observation using other variables or using `dcast()` putting the mean, variance, or sum to the value (`dcast()` is the function in `reshape2`)

Tyler Rinker's [minimal guide to `tidyr`](https://github.com/trinker/tidyr_in_a_nutshell).

Activity #2
  * Make a tibble with one row per year and columns for life expectancy for two or more countries.
    - Use `knitr::kable()` to make this table look pretty in your rendered homework.
    - Take advantage of this new data shape to scatterplot life expectancy for one country against that of another.
    
This tibble makes it easier to compare the life expectancy for the selected countries. `spread()` is useful for this situation. Also, is easier plotting a scatterplot values for one country against those of another. This is because all the countries' name becomes the header of each column.
I compared the life expectancy for Japan with that for Republic of Korea. I added reference line which is `y=x`. It helps to analyze and interpret the plot. The life expectancy for Japan is always higher than that for Republic of Korea. 

```{r}
g3 <- gapminder %>% 
  select(country, year, lifeExp) %>% 
  filter(country %in% c("Korea, Rep.", "Japan", "Canada", "Australia")) %>% 
  spread(key = "country", value = "lifeExp")
knitr::kable(g3, format = "markdown")

g3 %>% 
  ggplot(aes(x = `Korea, Rep.`, y = Japan, color = year)) +
  geom_point() +
  labs(title = "Life Expectancy for Korea and Japan") +
  coord_fixed(xlim = c(45, 85), ylim = c(45, 85)) +
  geom_abline()
```



Activity #3

  * Compute some measure of life expectancy (mean? median? min? max?) for all possible combinations of continent and year. Reshape that to have one row per year and one variable for each continent. Or the other way around: one row per continent and one variable per year.
    - Use `knitr::kable()` to make these tables look pretty in your rendered homework.
    - Is there a plot that is easier to make with the data in this shape versis the usual form? If so (or you think so), try it! Reflect.
    

I computed the `mean` of life expectancy for all possible combinations of continent and year by using `group_by()` and `summarize()`. Then, I reshaped it using `spread()`.
As explained in the previous Activity, this reshaped form would be useful to compare one continent's life expectancy with another's. This is because all the continents' name becomes the header of each column.
```{r}
g4 <- gapminder %>%
  select(country, continent, year, lifeExp) %>% 
  group_by(year, continent) %>% 
  summarize(lifeExp_mean = mean(lifeExp)) %>% 
  spread(key = "continent", value = "lifeExp_mean")
knitr::kable(g4, format = "markdown", digits = 2)

g4 %>% 
  ggplot(aes(x = Asia, y = Americas, color = year)) +
  geom_point() +
  geom_path() +
  geom_text(aes(label = year), color = "black",size = 3.5, hjust = 1) +
  labs(title = "The Average Life Expectancy for Europe and Americas") +
  coord_fixed() +
  geom_abline()

```



Activity #4

  * In [Window functions](http://stat545.com/block010_dplyr-end-single-table.html#window-functions), we formed a tibble with 24 rows: 2 per year, giving the country with both the lowest and highest life expectancy (in Asia). Take that table (or a similar one for all continents) and reshape it so you have one row per year or per year * continent combination.
  
I used the gapminder dataset and formed a dataset which shows the Asian countries with the highest life expectancy and the lowest in each year. I could keep the highest and lowest life expectancy observations in each year using `filter(min_rank(desc(lifeExp)) < 2 | min_rank(lifeExp) < 2)`. 
```{r}
my_gap <- gapminder%>%
  filter(continent == "Asia") %>%
  select(year, country, lifeExp) %>%
  group_by(year) %>%
  filter(min_rank(desc(lifeExp)) < 2 | min_rank(lifeExp) < 2) %>% 
  mutate(min_rank(desc(lifeExp))) %>% 
  arrange(year) %>% 
  select(-lifeExp) %>% 
  spread(key = "min_rank(desc(lifeExp))", value = "country") %>% 
  rename(Highest =  `1`, Lowest  =  `2`)
#  print(n = Inf)

knitr::kable(my_gap, format = "markdown")
```



### Join, merge, look up

__Problem__: You have two data sources and you need info from both in one new data object.

__Solution__: Perform a __join__, which borrows terminology from the database world, specifically SQL.

__Prompts__:

Activity #1

  * Create a second data frame, complementary to Gapminder. Join this with (part of) Gapminder using a `dplyr` join function and make some observations about the process and result. Explore the different types of joins. Examples of a second data frame you could build:
    - One row per country, a country variable and one or more variables with extra info, such as language spoken, NATO membership, national animal, or capitol city. If you really want to be helpful, you could attempt to make a pull request to resolve [this issue](https://github.com/jennybc/gapminder/issues/13), where I would like to bring ISO country codes into the gapminder package.
    - One row per continent, a continent variable and one or more variables with extra info, such as northern versus southern hemisphere.
    

I created a second data frame complementary to Gapminder, which contained country, capital_city, and Language. For simplicity, I used gapminder data after 2000.
```{r}
extra_info <- data.frame(country=c("Korea, Rep.", "Japan", "China", "Canada","Australia", "United States", "Vatican City"),
                     capital_city=c("Seoul", "Tokyo", "Beijing", "Ottawa","Canberra", "Washington, D.C.", "Vatican City"),
                     Language = c("Korean", "Japanese", "Standard Mandarin", "English/French","English", "English", "Italian"))
extra_info

gapminder_2000s <- gapminder %>% 
  filter(year >= 2000)
gapminder_2000s
```

As mentioned above, I did not specify a character vector of variables to join by using `by = "country"`. This was because there was only one variable in common. 

1. `semi_join()` creates a subset of the `gapminder_2000s`(which is on the left), which have the countries that are only found in the `extra_info`(which is on the right).

2. `inner_join()` creates the dataset which has all the variables in `gapminder_2000s` and `extra_info`. The dataset only keeps observations available in both datasets. 

3. `right_join()` creates the dataset which has all the variables in `gapminder_2000s` and `extra_info`. The dataset only keeps observations available in `extra_info`(which is on the right).

4 `left_join()` is same as `right_join()` but the dataset keeps observations available in `gapminder_2000s`(which is on the left)

5. `full_join()` creates the dataset which has all the variables in `gapminder_2000s` and `extra_info`. The dataset keeps all the observations available in both datasets.

- The difference between `semi_join` and `inner_join` is that `semi_join()` takes only the variables on the left. This can be verified by seeing the dimension of each dataset. Inner-joined dataset's dimension is (12x6)(6 is same as the number of variables in `gapminder_2000s` ). On the contrary, the other one's dimension is (12x8).

- Right_joined dataset has one more observation than inner-joined dataset. This is because `Vatican City` is not included in the `gapminder_2000s`. This also can be observed by comparing `left_join()` and `full_join()`


```{r}
#semi_join()
semi_join(gapminder_2000s, extra_info)

#inner_join()
inner_join(gapminder_2000s, extra_info)

#right_join()
right_join(gapminder_2000s, extra_info)

#left_join()
left_join(gapminder_2000s, extra_info)

#full_join()
full_join(gapminder_2000s, extra_info)
```


